---
title: "Dynamic Foraging Patterns in Antarctic Humpbacks"
author: "Ben Weinstein"
date: "`r Sys.time()`"
output: 
  html_document:
    keep_md: yes
    number_sections: yes
    theme: spacelab
    toc: yes
---

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(knitr)
opts_chunk$set(echo=F,warning=F,message=F,fig.width = 11,fig.height = 5,cache=F)
library(boot)
library(raster)
library(tidyr)
library(ggplot2)
library(MASS)
library(ggmap)
library(dplyr)
library(chron)
library(gridExtra)
library(stringr)
library(R2jags)
library(maptools)
library(reshape2)
#New model to be run, flag turned off if just updating.
newModel<-F

##ggplot theme
mytheme<-theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),axis.title.x=element_blank(),axis.title.y=element_blank(),panel.grid=element_blank())

```

```{r,eval=T}
#Load in data
load("DynamicForaging.RData")
#jags$recompile()
newModel<-F
```

```{r}
#read data
mdat<-read.csv("InputData/Antarctic Humpback overlap with krill fisheries -3085052751565071394.csv")
#standardize column names to match the simulation
#Create an animal tag.
mxy <- as(mdat, "data.frame")
mxy$Animal<-mxy$individual.local.identifier
mxy$x<-mxy$location.long
mxy$y<-mxy$location.lat

#remove minke whales, they should already be gone.
minke<-c("131117","131118","131120","154184")
mxy<-mxy[!mxy$individual.local.identifier %in% minke,]

#grab set of animals
#mxy<-mxy[mxy$Animal %in% c("131143","131142"),]

#empty coordinates
mxy<-mxy[!is.na(mxy$x),]

#crop by extent
d<-SpatialPointsDataFrame(cbind(mxy$x,mxy$y),data=mxy,proj4string=CRS("+proj=longlat +datum=WGS84"))

cropoly<-readShapePoly("InputData/CutPolygon.shp",proj4string=CRS("+proj=longlat +datum=WGS84"))

b<-d[!is.na(d %over% cropoly)[,2],]

mxy<-b@data

#remove migration events, create user cut polygon

#set datestamp
mxy$timestamp<-as.POSIXct(mxy$timestamp,format="%Y-%m-%d %H:%M:%S.000")

#month and year columns
mxy$Month<-months(mxy$timestamp)
mxy$Year<-years(mxy$timestamp)

#Only austral sping and summer
mxy<-mxy[mxy$Month %in% month.name[1:5],]

#remove empty timestamps
mxy<-mxy[!is.na(mxy$timestamp),]

#remove duplicates
mxy<-mxy[!duplicated(data.frame(mxy$timestamp,mxy$Animal)),]

```

```{r}
#hold copy of orig
oxy<-mxy

#remake spatial object
d<-SpatialPointsDataFrame(cbind(mxy$x,mxy$y),data=mxy,proj4string=CRS("+proj=longlat +datum=WGS84"))

#Plot
pp<-c(mean(mxy$x),mean(mxy$y))
m <- get_map(location=pp,source="google",zoom=4,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal)),size=.5) + scale_color_discrete(guide="none") + mytheme
```

##By Month

```{r,fig.height=12}
mxy$Month<-factor(mxy$Month,levels=month.name)
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal)),size=.5)  + facet_wrap(~Month,nrow=2) + scale_color_discrete(guide="none") + theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank()) + labs(x="",y="")
ggsave("Figures/RawTracks.jpg",dpi=300,height=5,width=7)

#black and white
ggplot()+geom_path(data=mxy, aes(x=x, y=y,group=Animal),size=.5)  + facet_wrap(~Month,nrow = 2)  + borders(fill="grey90") + coord_cartesian(ylim = c(-70,-62),xlim=c(-55,-73)) + theme_bw() + mytheme + scale_fill_manual(values=rep("NA",8)) + scale_color_discrete(guide="none")
ggsave("Figures/MonthTracks_BW.jpeg",height=4, width=5)
```

#Correlated random walk

*Process Model*

$$ d_{t} \sim T*d_{t-1} + Normal(0,\Sigma)$$
$$ x_t = x_{t-1} + d_{t} $$

## Parameters

For each individual:

$$\theta = \text{Mean turning angle}$$
$$\gamma = \text{Move persistence} $$

For both behaviors process variance is:
$$ \sigma_{latitude} = 0.1$$
$$ \sigma_{longitude} = 0.1$$

##Behavioral States

$$ \text{For each individual i}$$
$$ Behavior_1 = \text{traveling}$$
$$ Behavior_2 = \text{foraging}$$

$$ \alpha_{i,1,1} = \text{Probability of remaining traveling when traveling}$$
$$\alpha_{i,2,1} = \text{Probability of switching from Foraging to traveling}$$

$$\begin{matrix}
  \alpha_{i,1,1} & 1-\alpha_{i,1,1} \\
  \alpha_{i,2,1} & 1-\alpha_{i,2,1} \\
\end{matrix}$$

With the probability of switching states:

$$logit(\phi_{traveling}) = \alpha_{Behavior_{t-1}}$$

$$\phi_{foraging} = 1 - \phi_{traveling} $$

##Continious tracks

The transmitter will often go dark for 10 to 12 hours, due to weather, right in the middle of an otherwise good track. The model requires regular intervals to estimate the turning angles and temporal autocorrelation. As a track hits one of these walls, call it the end of a track, and begin a new track once the weather improves. We can remove any micro-tracks that are less than three days.
Specify a duration, calculate the number of tracks and the number of removed points. Iteratively.

```{r}
##Time is the beginning of the first point.
step_length=4

sxy<-split(mxy,mxy$Animal)

#time diff function
timed<-function(d,step_length){
  d$j[1]<-0
  for (x in 2:nrow(d)){
    d$j[x]<-as.numeric(difftime(as.POSIXct(d$timestamp[x]),as.POSIXct(d$timestamp[x-1]),units="mins"))/(step_length*60)
  }
  
  #Split out track endings
  ends<-c(1,which(d$j>1),nrow(d))

  for(w in 2:length(ends)){
    d[ends[w-1]:ends[w],"Track"]<-w-1
  }
  
  #remove tracks that are shorter than three days
  track_time<-d %>% group_by(Track) %>% summarize(mt=difftime(max(as.POSIXct(timestamp)),min(as.POSIXct(timestamp)),units="hours")) %>% filter(mt>=24) %>% .$Track
  
  d<-d[d$Track %in% track_time,]
  
  #renumber the tracks
  d$Track<-as.numeric(as.factor(d$Track))
  return(d)
  }

sxy<-lapply(sxy,timed,step_length=step_length)

#Format matrices for jags
mxy<-rbind_all(sxy)

######recode whales
#mxy$Animal<-as.numeric(as.factor(mxy$Animal))

sxy<-split(mxy,list(mxy$Animal,mxy$Track),drop=TRUE)

sxy<-lapply(sxy,function(x){
#How many observations in each step length segment
x$step<-as.numeric(cut(as.POSIXct(x$timestamp),paste(step_length,"hours")))
return(x)
})

mxy<-rbind_all(sxy)

#refactor animal
mxy$Animal<-as.numeric(as.factor(mxy$Animal))
```

### After filitering

```{r,fig.height=12}
mxy$Month<-factor(mxy$Month,levels=month.name)
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal)),size=.5)  + facet_wrap(~Month,nrow=2) + scale_color_discrete(guide="none") + theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank()) + labs(x="",y="")
ggsave("Figures/RawTracks.jpg",dpi=300,height=5,width=7)
```

```{r}
#total number of steps per track/animal
steps_all<-mxy %>% group_by(Animal,Track) %>% summarize(n=length(unique(step)))

# give each step a label
mxy<-mxy %>% group_by(Animal,Track,step) %>% mutate(jStep=1:n())

#Cast time array
j<-acast(mxy,Animal~Track~step~jStep,value.var="j")

#how many observations per individual in each step
mxy$step<-factor(mxy$step,levels=1:max(steps_all$n))
idx<-melt(table(mxy$Animal,mxy$Track,mxy$step))
colnames(idx)<-c("Animal","Track","step","jStep")
idx<-acast(data=idx,Animal~Track~step)

#month array
mxy$MonthF<-as.numeric(factor(mxy$Month,levels=month.name))

MonthA<-acast(mxy,Animal~Track~step,value.var="MonthF",fun.aggregate = min)
MonthA[!is.finite(MonthA)]<-NA

#Individuals
ind=length(unique(mxy$Animal))

#tracks per indivudal
tracks<-mxy %>% group_by(Animal) %>% summarize(tracks=length(unique(Track))) %>% .$tracks

#steps per track
steps<-acast(steps_all,Animal~Track,value.var="n")

#index array for merging
mxy$Index<-1:nrow(mxy)

#obs array
obs<-melt(mxy,measure.vars=c("x","y"))
obs<-acast(obs,Animal~Track~step~jStep~variable)

MonthIndex<-acast(mxy,Animal~Track~step,value.var="Month",fun=function(x) names(which.max(table(x))) )
```

How did the filter change the extent of tracks?

```{r}
#Plot
mxy<-mxy %>% arrange(Animal,Track,step,jStep)

pp<-c(mean(mxy$x),mean(mxy$y))
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal),group=paste(Animal,Track)),size=.5) + scale_color_discrete(guide="none") + mytheme
```

```{r,fig.height=15}
ggplot() +  geom_path(data=oxy,col='blue',aes(x=x,y=y)) +  geom_path(data=mxy,col='red',aes(x=x,y=y),size=2,alpha=0.7)  + facet_wrap(~individual.local.identifier,scales="free",ncol=4)
```

```{r,fig.height=3,fig.width=5}
ggplot(data=steps_all,aes(x=n*step_length/24)) + geom_histogram() + labs(x="Days") + ggtitle("Track Length")
ggplot(data=steps_all,aes(x=Track)) + geom_histogram() + labs(x="Subtracks per Animal")
```

```{r,child="Bayesian/MultiSpecies.R",eval=T}
```

```{r,eval=T}
#source jags file
source("Bayesian/MultiSpecies.R")

#prior cov shape
R <- diag(c(1,1))
data=list(argos=obs,steps=steps,R=R,ind=ind,j=j,idx=idx,tracks=tracks,Month=MonthA,Months=max(MonthA,na.rm=T))

#paramters to track
pt<-c("theta","gamma","phi","alpha_mu","state")

if(newModel){
  system.time(jagM<-jags.parallel(model.file = "Bayesian/Multi_RW.jags",data=data,n.chains=2,parameters.to.save=pt,n.iter=40000,n.burnin=39000,n.thin=5,DIC=FALSE))
}

```

```{r,eval=F}
#update function
recompile(jagM)
jagM <- update(jagM,n.iter=300,n.burnin=0,n.thin=0) 
```

##Chains
```{r,eval=newModel}

#delete jags objects
#rm(data)
#rm(argos)
#rm(obs)
#rm(j)
gc()

#bind chains
pc<-melt(jagM$BUGSoutput$sims.array)

rm(jagM)
gc()

colnames(pc)<-c("Draw","chain","par","value")

#extract parameter name
pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]

#Extract index
splitpc<-split(pc,pc$parameter)

#single index
splitpc[c("theta")]<-lapply(splitpc[c("theta")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,Behavior=sv)
    return(pc)
})

##Double index
#single index
splitpc[c("alpha_mu","gamma")]<-lapply(splitpc[c("alpha_mu","gamma")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+)]"))[,3:4]
    colnames(sv)<-c("Behavior","Month")
    pc<-data.frame(x,sv)
    return(pc)
})

#Three index
splitpc[c("phi")]<-lapply(splitpc[c("phi")],function(x){
#As matrices
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+),(\\d+)]"))[,3:6]
    colnames(sv)<-c("Animal","Track","step","Behavior")
    pc<-data.frame(x,sv)
})

#State index
splitpc[c("state")]<-lapply(splitpc[c("state")],function(x){
#As matrices
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+)]"))[,3:5]
    colnames(sv)<-c("Animal","Track","step")
    pc<-data.frame(x,sv)
})

#bind all matrices back together
pc<-rbind_all(splitpc)
rm(splitpc)

#associate month with state
mm<-melt(MonthIndex)
colnames(mm)<-c("Animal","Track","step","MonthA")
pc<-merge(pc,mm,by=c("Animal","Track","step"),all.x=T)

#complete level matching, ugly but successful

pc[pc$parameter %in% c("alpha_mu","gamma"),]$MonthA<-month.name[as.numeric(pc$Month[pc$parameter %in% c("alpha_mu","gamma")])]
pc$MonthA<-factor(pc$MonthA,levels=month.name)

```

```{r,fig.height=10}
#plot all but phi and state
ggplot(pc[!pc$parameter %in% c("phi","state"),],aes(x=Draw,y=value)) + facet_wrap(~par,scales="free",ncol=3) + geom_line(aes(col=as.factor(chain))) + labs(col="Chain")

a<-pc %>% filter(parameter %in% c("state","phi")) %>% filter((parameter == "phi"&Behavior==2)|parameter =="state") %>% dplyr::select(Draw,chain,Animal,Track,step,value,parameter) %>% spread(key=parameter,value=value)
ggplot(a,aes(x=phi,fill=as.factor(state-1))) + geom_density(alpha=0.7) + labs(fill="Behavior")
```

```{r,eval=F}
#write.csv
#write.csv(pc,"OutData/pars.csv")
#pc<-read.csv("OutData/pars.csv")
pc$Behavior<-as.factor(pc$Behavior)
```

```{r,fig.height=20}
#Plot
ggplot(data=pc[!pc$parameter %in% c("phi","state"),],aes(x=value)) + geom_histogram() + facet_wrap(~par,scales='free',ncol=3) + theme_bw()  + ggtitle("Estimated parameters")
```

## Change in autocorrelation over time

```{r}
g<-pc[pc$parameter %in% c("gamma"),] %>% mutate(bstate=ifelse(Behavior==1,"Traveling","Area-restricted Search"))
ggplot(data=g,aes(x=value,fill=MonthA)) + geom_density(alpha=0.7) + theme_bw()+ facet_wrap(~bstate,scales="free") + labs(fill="Month",x="Degree of movement autocorrelation") + scale_fill_brewer(palette="RdBu")

ggsave("Figures/Gammaplot.jpeg",height=3,width=6,dpi=300)

a<-pc[pc$parameter %in% c("gamma"),] %>% mutate(bstate=ifelse(Behavior==1,"Traveling ","Foraging"))

ggplot(data=a,aes(y=value,x=MonthA)) + geom_boxplot() + theme_bw()  + ggtitle("Movement autocorrelation") + labs(x="Month",fill="Behavior") + facet_wrap(~bstate,scales="free")
ggsave("Figures/gammabox.jpeg",height=3,width=8)

```

# Change in transition probabilities over time

```{r}
a<-pc[pc$parameter %in% c("alpha_mu"),] %>% mutate(bstate=ifelse(Behavior==1,"Remain Traveling ","Switch to Traveling"))
ggplot(data=a,aes(x=inv.logit(value),fill=MonthA)) + geom_density(alpha=0.7) + theme_bw()+ facet_wrap(~bstate,scales="free") + labs(fill="Month",x="Transition Probability") + scale_fill_brewer(palette="RdBu")
ggsave("Figures/Alphaplot.jpeg",height=4,width=6,dpi=300)

ggplot(data=a,aes(y=inv.logit(value),x=MonthA)) + geom_boxplot() + theme_bw()+ facet_wrap(~bstate,scales="free") + labs(x="Month",y="Transition Probability")
ggsave("Figures/AlphaBox.jpeg",height=3,width=8)
```

## Parameter Summary

```{r,fig.height=10,fig.width=15}
psum<-pc %>% filter(!parameter %in% c("phi","state")) %>% group_by(parameter,par) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

as.data.frame(psum)

ggplot(psum,aes(x=par,y=mean,ymax=upper,ymin=lower)) + geom_linerange() + geom_point() + facet_wrap(~parameter,scales="free",ncol=2)

#format gamma
gammadf<-pc %>% filter(parameter=="gamma") %>% group_by(MonthA,Behavior) %>% summarize(mean=mean(value))

#format alpha
alphadf<-pc %>% filter(parameter=="alpha_mu") %>% group_by(MonthA,Behavior) %>% summarize(mean=mean(value))

#both
simdf<-pc %>% filter(parameter %in% c("gamma","alpha_mu")) %>% group_by(MonthA,Behavior,parameter) %>% summarize(mean=mean(value))

```

#Behavioral Prediction

```{r}
#mean transition probability
mean_phi<-pc %>% filter(parameter=="phi",Behavior==2) %>% mutate(step=as.numeric(as.character(step))) %>% group_by(Animal,Track,step) %>% summarize(phi=median(value)) %>% arrange(Animal,Track,step)

#mode state for each position
mode_state<-pc %>% filter(parameter %in% "state") %>% mutate(bstate=ifelse(value==1,"Traveling","Area-restricted Search")) %>% group_by(Animal,Track,step,bstate) %>% summarize(n=n()) %>% spread(key=bstate,value=n,fill=0)%>% mutate(ratio=`Area-restricted Search`/Traveling) %>% mutate(state=as.character(ifelse(ratio>=1,"Area-restricted Search","Traveling")))

mxy<-merge(mxy,mean_phi,by=c("step","Track","Animal"))
mxy<-merge(mxy,mode_state,by=c("step","Track","Animal"))

#make factor
mxy$bstate<-factor(mxy$state)

#make phi a factor
mxy$phistate<-cut(mxy$phi,c(0,0.5,1),labels=c("Traveling","Area-restricted Search"))

#ensure order for plotting
mxy<-mxy %>% arrange(Animal,Track,step,jStep)
```

```{r}
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color="bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=bstate,group=paste(Animal,Track)),size=.5,alpha=0.7)   + mytheme
```

##Spatial Prediction

```{r,fig.width=9}
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color="bw")

ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=bstate,group=paste(Animal,Track)),size=.2) + mytheme + labs(col="Behavior") +  facet_wrap(~Month,nrow=2)

ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=phistate,group=paste(Animal,Track)),size=.2)  + facet_wrap(~Month,nrow=2)
ggsave("Figures/SpatialPred.jpg",height=5,width=5)

#Black and white version
ggplot()+geom_path(data=mxy, aes(x=x, y=y,col=bstate,group=paste(Animal,Track)),size=1)  + borders(fill=NA,alpha=0.9,size=0.5) + coord_cartesian(ylim = c(-69,-62),xlim=c(-55,-72)) + theme_bw() + mytheme + scale_fill_manual(values=rep("NA",8)) + labs(col="Behavior") + scale_color_manual(values=c("grey10","grey60"))  + facet_wrap(~Month,nrow=1) + theme(legend.position="bottom")
ggsave("Figures/SpatialPred_BW.jpg",height=3,width=10)

```

## By individual

```{r,fig.height=12}
ggplot()+geom_path(data=mxy, aes(x=x, y=y,col=bstate,group=paste(Animal,Track)),size=.5)   + mytheme + facet_wrap(~Animal,scales="free")
```

##Autocorrelation in behavior

```{r,fig.height=25,fig.width=11}
#create proper time stamp
ggplot(data=mxy,aes(x=as.POSIXct(timestamp),y=as.factor(state))) + geom_line(aes(group=Track),size=1)+ theme_bw()  + facet_wrap(~individual.local.identifier,ncol=2,scales="free") + labs(x="Time",y="Behavior") 
```

#Simulated tracks

```{r,fig.height=8,fig.width=12}
source("Sim.R")

#format gamma df
gt<-simdf %>% spread(key=Behavior,value=mean) %>% droplevels()
gt<-split(gt[,2:4],gt$MonthA)

#monthly trajectories
mt<-lapply(gt,function(k){
  
  mxy<-list()

  #individuals
  ind<-2
  
  #Total time length for each individual track
  step.ind<-rep(50,4)
  
  #time for each step
  step_length=4

  for(x in 1:ind){
    
    #autocorrelation is the second row
    #transition is first row, needs to be logit transformed.
      mxy[[x]]<-traj(gamma=as.numeric(k[2,2:3]),theta=c(6,0),a1=inv.logit(as.numeric(k[1,2:3])),total_time = step.ind[x],step_length=step_length)
  } 
  return(mxy)
})

mt<-melt(mt,id.vars=colnames(mt[[1]][[1]]))

#Label Individual and month
colnames(mt)[colnames(mt) %in% 'L1']<-"Month"
colnames(mt)[colnames(mt) %in% 'L2']<-"Animal"

mt$Month<-factor(mt$Month,levels=month.name)
ggplot(mt,aes(x=x,y=y)) + theme_bw()  + geom_path(aes(col=State,group=Animal),size=1)+ facet_wrap(~Month,scales='free')  + theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank()) + labs(x="",y="")
ggsave("Figures/SimulatedTracks.jpg",dpi=300,height=4,width=7)
```

##Behavioral description

## Predicted behavior duration

```{r,fig.height=5,fig.width=14}

#associate Month with state change
runf<-function(x){
  

  #arrange by step
  x$step<-as.numeric(x$step)
  x<-x %>% arrange(step)
  
  #calculate run length
  state_change<-rle(as.numeric(x$value))
  runl<-data.frame(Animal=unique(x$Animal),runs=state_change$lengths,Behavior=state_change$values)
  runl$Behavior<-as.factor(runl$Behavior)
  levels(runl$Behavior)<-c("Traveling","Foraging")
  
  return(runl)
}

runs<-pc %>% filter(parameter=='state') %>% group_by(Animal,Track,Draw,chain) %>% do(runf(.))

#turn steps to hours
runs$hours<-runs$runs*step_length
```

```{r,fig.height=4}
mr<-runs %>% group_by(Behavior) %>% summarize(mean=mean(hours),lower=quantile(hours,0.05),upper=quantile(hours,0.95))

mr$bstate<-as.factor(mr$Behavior)
levels(mr$bstate)=c("Traveling","Foraging")

ggplot(mr,aes(col=bstate,y=mean,ymin=lower,ymax=upper,x=Month)) + geom_pointrange(size=1.5,position = position_jitter(width = .3)) + theme_bw() + ggtitle("Average Duration of Behavior") + labs(y="Hours")
ggsave("Figures/Duration.jpeg",height=4,width=7)
```

## Duration by month

```{r}
runs<-pc %>% filter(parameter=='state') %>% group_by(Animal,Track,Draw,chain,MonthA) %>% do(runf(.))

#turn steps to hours
runs$hours<-runs$runs*step_length

#ignore single blips.
mr<-runs %>% group_by(Behavior,MonthA) %>% filter(!runs %in% 1:2) %>% summarize(mean=mean(hours),lower=quantile(hours,0.05),upper=quantile(hours,0.95)) 
mr$bstate<-as.factor(mr$Behavior)
levels(mr$bstate)=c("Traveling","Foraging")

ggplot(mr,aes(x=MonthA,y=mean,ymin=lower,ymax=upper)) + geom_pointrange(size=1.2,alpha=.7) + theme_bw() + labs(y="Duration of Behavior (Hours)",x="Month") + facet_wrap(~bstate,scales="free")
ggsave("Figures/Duration.jpeg",height=4,width=6,dpi=300)

ggplot(runs,aes(x=MonthA,y=hours)) + geom_boxplot() + theme_bw() + ggtitle("Average Duration of Behavior") + labs(y="Hours",col="Month",x="Behavior") + facet_wrap(~Behavior,scales="free")
```

```{r}
#Every run eventually ends in a switch (that's what makes it a run)
runs$status<-1

endtrack<-function(x){
  x[nrow(x),"status"]<-0
  return(x)
}

#if track ends, status is 0 not 1
feedr<-runs %>% group_by(Draw,chain,Animal) %>% do(endtrack(.)) %>% filter(Behavior=="Foraging")

#save runs to file
#write.csv(feedr,"OutData/Runs.csv")
```

## Duration by month

```{r}
runs<-mxy %>% group_by(Animal,Track,Month) %>% mutate(value=phistate) %>% do(runf(.))

#turn steps to hours
runs$hours<-runs$runs*step_length

mr<-runs %>% group_by(Behavior,Month) %>% summarize(mean=mean(hours),lower=quantile(hours,0.05),upper=quantile(hours,0.95)) 
mr$bstate<-as.factor(mr$Behavior)
levels(mr$bstate)=c("Traveling","Foraging")

ggplot(mr,aes(x=Month,y=mean,ymin=lower,ymax=upper)) + geom_pointrange(size=1.2,alpha=.7) + theme_bw() + labs(y="Duration of Behavior (Hours)",x="Month") + facet_wrap(~bstate,scales="free")
ggsave("Figures/Duration.jpeg",height=4,width=6,dpi=300)

ggplot(runs,aes(x=Month,y=hours)) + geom_boxplot() + theme_bw() + ggtitle("Average Duration of Behavior") + labs(y="Hours",col="Month",x="Behavior") + facet_wrap(~Behavior,scales="free")
```

```{r}
#Every run eventually ends in a switch (that's what makes it a run)
runs$status<-1

endtrack<-function(x){
  x[nrow(x),"status"]<-0
  return(x)
}

#if track ends, status is 0 not 1
feedr<-runs %>% group_by(Draw,chain,Animal) %>% do(endtrack(.)) %>% filter(Behavior=="Foraging")

#save runs to file
#write.csv(feedr,"OutData/Runs.csv")
```

#Proportion of states by month

```{r}
prop<-pc %>% filter(parameter=="state") %>% group_by(MonthA,Draw,chain,value) %>% summarize(s=n()) 
ratio<-pc %>% filter(parameter=="state") %>% group_by(MonthA,Draw,chain,value) %>% summarize(s=n()) %>% spread(key=value,value=s) %>% mutate(b=(`2`/(`1`+`2`))) 

ggplot(prop,aes(fill=as.factor(value),y=s,x=MonthA)) + geom_violin() + labs(x="Month",y="Hours",fill="Behavior") + scale_fill_discrete(label=c("Traveling","Area-restricted Search")) + theme_bw()

ggplot(ratio,aes(y=b,x=MonthA)) + geom_boxplot() + labs(y="Proportion Area-restricted Search",x="Month") + theme_bw()  + scale_fill_brewer(palette="RdBu") + scale_y_continuous(labels = scales::percent)
ggsave("Figures/PropForaging.jpeg",height=3,width=5)
```

#Time between foraging bouts

```{r}
timebetween<-function(a){
  
  a<-a %>% arrange(step)
  
  a$timestamp<-as.POSIXct(a$timestamp,format="%Y-%m-%d %I:%M:%S %p")
  
  #run length encoding
  ra<-rle(as.numeric(a$phistate))
  
  #create index matrix
  rindex<-data.frame(Behavior=ra$values,Runs=ra$lengths,Index=cumsum(ra$lengths))
  
  if(nrow(rindex)<2){return(data.frame())}
  for (x in 2:nrow(rindex)){
    rindex[x,"Hours"]<-difftime(a[[rindex$Index[x],"timestamp"]],a[[rindex$Index[x-1],"timestamp"]],units="hours")
  }

  out<-rindex %>% filter(Behavior==2)
  return(out) 
  }

#calculate time between
tb<-mxy %>% group_by(Animal,Track,Month)  %>% do(timebetween(.)) %>% filter(!is.na(Hours))

ggplot(tb,aes(x=Month,y=Hours)) + geom_boxplot() + theme_bw() + labs(x="Month",y="Time between foraging bouts (Hours)")

ggplot(tb,aes(x=Hours,fill=Month)) + geom_density(alpha=0.4) + xlab("Time Between Foraging Bouts (Hours)") + theme_bw() + scale_fill_brewer(palette = "RdBu")

mean_bout<-tb %>% group_by(Month) %>% summarize(m=median(Hours))
ggplot(mean_bout,aes(x=Month,y=m)) + geom_boxplot() + theme_bw() + labs(x="Month",y="Average time between bouts")
ggsave("Figures/TimeBetweenBouts.jpeg",height=4,width=6)
```

#Distance between bouts

```{r}
#function for calculating geographic distance between bouts
distbout<-function(a){
  
  a<-a %>% arrange(Animal,Track,step)
  
  stater<-rle(as.numeric(a$phistate))
  Index<-cumsum(stater$lengths[-length(stater$lengths)])
  
  if(length(Index) <= 1){distframe<-data.frame(NULL)} else {
  out<-matrix(ncol=1,nrow=length(Index)-1)
  for(x in 1:(length(Index)-1)){
    pts<-mxy[c(Index[x],Index[x+1]+1),]
    
    pts<-SpatialPoints(cbind(pts$x,pts$y),CRS('+proj=longlat +ellps=WGS84'))
    out[x]<-spDists(pts,pts,longlat=T)[1,2]
  }
  
  distframe<-data.frame(Animal=unique(a$Animal),Track=unique(a$Track),Month=min(as.numeric(unique(a$Month))),km=out)
  }
  return(distframe)
  }

#calculate distance between foraging bouts for all tracks
distf<-mxy %>% group_by(Animal,Track) %>% do(distbout(.)) %>% mutate(Month=month.name[Month])

#average distance by month
avg_dist<-distf %>% group_by(Month) %>% summarize(meankm=median(km),lower=quantile(km,0.05),upper=quantile(km,0.95))

avg_dist$Month<-factor(avg_dist$Month,levels=month.name)

ggplot(avg_dist,aes(x=Month,y=meankm)) + geom_pointrange(aes(y=meankm,ymin=lower,ymax=upper)) + theme_bw() + labs(y="Average distance between consecutive foraging bouts")
```

#Time spent in grid cell

```{r,fig.height=7,fig.width=11}

#time in each cell function
  timecell<-function(j){
    
    rlej<-rle(x=diff(j$Index))
    position<-c(1,cumsum(rlej$lengths))
    timestack<-c()
    
    #if only one position
    if(length(position)==1){
      timestack[1]<-0
      
    } else {
      
      #time difference between first and last point in the cell
        for(x in 2:length(position)){
          timed<-j[position[x-1]:position[x],]
          timestack[x-1]<-difftime(max(timed$timestamp),min(timed$timestamp),units="days")
      }
    }
  
    out<-data.frame(Sum=sum(timestack))
    return(out)
  }
  
restime<-function(mxy){
  #raster
  r<-raster(extent(d)*c(1.1,1.1),nrow=100,ncol=100)
  
  #res(r)<-0.25
  #put cellindex number
  r[]<-1:ncell(r)
  
  sxy<-split(mxy,mxy$Animal)
  sxy<-lapply(sxy,function(x) {x$Index<-1:nrow(x); return(x)})
  
  
  #loop through all animals
  sptracks<-lapply(sxy,function(x){
    k<-SpatialPointsDataFrame(cbind(x$x,x$y),data=x,proj4string=CRS("+proj=longlat +datum=WGS84"))
    
    #get the cells that points are in
    k$cellvalue<-raster::extract(x=r,y=k)
    
    #for each cell value, perform time function
    sptrack<-as.data.frame(k) %>% group_by(Animal,cellvalue) %>% do(timecell(.))
    return(sptrack)}
  )
  
  sptracks<-rbind_all(sptracks)
  
  #get sum across animals
  cellvals<-sptracks %>% group_by(cellvalue) %>% summarize(Sum=sum(Sum))
  
  #is there a better way to loop through?
  r[]<-NA
  for(x in 1:nrow(cellvals)){
    r[cellvals$cellvalue[x]]<-cellvals$Sum[x]
  }
  return(r)
}

plot(restime(mxy))

#split tracks by month data, just get foraging data
md<-split(mxy[mxy$bstate %in% "Area-restricted Search",],mxy[mxy$bstate %in% "Area-restricted Search","Month"],drop=T)

r<-stack(lapply(md,restime))

plot(r)

#turn to 
f<-melt(data.frame(rasterToPoints(r)),id.vars=c("x","y"))

#remove NA's
f<-f %>% filter(!is.na(value))

ggmap(m) + geom_tile(data=f,aes(x=x,y=y,fill=value))  + scale_fill_continuous(low="blue",high="red") + labs(fill="Days")  + facet_wrap(~variable,nrow=1)  + theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank()) + labs(x="",y="") + theme(legend.position="bottom")
ggsave("Figures/ResidencyMap.jpg",height=5,width=9,dpi=300)
```

```{r}
save.image("DynamicForaging.RData")
```